---
title: "Model mismatch"
date: "`r Sys.Date()`"
output:
  bookdown::html_document2:
    code_folding: show
    code_download: true
    theme: sandstone
    toc: true
    toc_float: true
    number_sections: true
    fig_caption: yes
---

```{r setup, message=FALSE}
# Data manipulation
library(tidyverse)
library(data.table) 

# Spatial operations
library(sf)
library(raster)
library(stars)

# Matrix operations
library(Matrix)

# MNO data handling and propagation model setup
# Credits to Prof. Martijn Tennekes https://github.com/mtennekes/mobloc
library(mobloc)

# Comparison of 2d histograms (Kantorovitch Wasserstein distance a.k.a. Earth Movers distance)
# Credits to Prof. Stefano Gualandi https://cran.r-project.org/web/packages/SpatialKWD/SpatialKWD.pdf
library(SpatialKWD)

# Output organisation and plotting support
library(ggthemes)
library(viridis)
library(ggrepel)
library(ggpointdensity)
library(scattermore)
library(grid)
library(gridExtra)
library(knitr)
library(DT)

# seed for reproducibility
set.seed(42)


# Loading Custom functions
source("pipeline functions.R")

# Load toyworld objects 
area <- readRDS("workflow_objects/area.rds")
cellplans.list <- readRDS("workflow_objects/cellplans.list.rds")
gen.model.objects <- readRDS("workflow_objects/gen.model.objects.rds")
```

## Model mismatch techniques

### Random noise implementation

```{r noise-implementation}
# define the number of bits --> final number of levels are 2^n
noise.expression <- c(no.03 = "runif(min = -3, max = 3",
                      no.06 = "runif(min = -6, max = 6",
                      no.09 = "runif(min = -9, max = 9",
                      no.12 = "runif(min = -12, max = 12",
                      no.15 = "runif(min = -15, max = 15",
                      no.18 = "runif(min = -18, max = 18",
                      no.21 = "runif(min = -21, max = 21",
                      no.true = "runif(min = 0, max = 0")

noise.expression.wo.true <- names(noise.expression)[!names(noise.expression) == ("no.true")]
noise.expression.only.true <- names(noise.expression)[names(noise.expression) == ("no.true")]


noise_mag <- function(x, expression, digits) {
  
  n <- length(x)
  
  expression.comp <- paste0(expression, ", n = ", n, ")")
    
  amount <- eval(parse(text = expression.comp))
  
  final <- x + round(amount, digits = digits)
  
  return(final)
  
  
}

noise_mutate <- function(df, noise.expression, digits) {
  
  mutate(df, "dBm.noise.{noise.expression}" := noise_mag(x = dBm, expression = noise.expression, digits = digits))
}


P.long.noise.df.helper <- cellplans.list %>% 
  map(~mutate(.x$signal.strength.comb.dt, cell.kind = substr(cell, 1, 2))) %>% 
  map2(., cellplans.list, ~left_join(.x, .y$param.df, by = "cell.kind")) %>% 
  map_dfr(~dplyr::rename(., sig_d_th = dominance.th), .id = "network.kind") %>% 
  group_by(network.kind)

set.seed(222)

P.long.noise.complete.df <- noise.expression %>% 
  map(~noise_mutate(df = P.long.noise.df.helper, noise.expression = .x, digits = 2)) %>% 
  map(~ungroup(.)) %>% 
  map2(., names(noise.expression), ~rename(.x, !!.y := starts_with("dBm.noise."))) %>% 
  map_at(noise.expression.wo.true, ~dplyr::select(., starts_with("no."))) %>%
  bind_cols() %>% 
  pivot_longer(cols = starts_with("no."), names_to = "noise.level", values_to = "noise.mag") %>% 
  split(list(.$network.kind, .$noise.level)) %>% 
  map(~mutate(., s.mismatch.mag = db2s(noise.mag, midpoint, steepness)))

names.network.noise <- names(P.long.noise.complete.df)

P.long.noise.df <- P.long.noise.complete.df %>% 
  map(~dplyr::select(., network.kind, matches("tile.id|cell"), noise.level, s.mismatch.mag, sig_d_th)) %>% 
  map(~left_join(., area$area.prior, by = "tile.id.chr")) %>% 
  map(~filter(., !s.mismatch.mag < 0.05)) %>% # define minimum threshold 
  map(~mutate(., pij = con_llh_sens_custom(., "s.mismatch.mag", digits = 3)))

# saveRDS(P.long.noise.df, "workflow_objects/P.long.noise.df.rds")  
P.long.noise.df <- readRDS("results/P.long.noise.df.rds")
```


```{r}
# noise on the dBm scale
order.levels.noise <- c(noise.expression.only.true, noise.expression.wo.true)
dense.names.noise <- names(P.long.noise.df) %>% 
  str_subset("cellplan.3") 

(dBm.plot.ME <- P.long.noise.complete.df %>%
  magrittr::extract(dense.names.noise) %>% 
  map_dfr(~dplyr::select(., cell, cell.kind, tile.id.num, noise.mag, noise.level)) %>% 
  filter(cell == ("ME.123.C.2")) %>% 
  mutate(noise.level = factor(noise.level, levels = order.levels.noise)) %>% 
  left_join(area$area.sf, by = "tile.id.num") %>% 
  st_as_sf() %>% 
  ggplot() +
  geom_sf(aes(fill = noise.mag), color = "transparent") +
  scale_fill_viridis(na.value = "transparent") +
  facet_wrap(vars(noise.level), ncol = 3) +
  coord_sf() +
  labs(title = "Meso noise on the dBm scale without minimum threshold", 
       color = "dBm", fill = "dBm", x = "", y = "") +
  theme(axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.x = element_blank(),
        axis.ticks.y = element_blank(),
        legend.position = "bottom"))
# ggsave("workflow_objects/noise.dBm.plot.ME.png", dBm.plot.ME, device = "png")
```

```{r}
# noise on the signal dominance scale

# paper.filter.noise <- order.levels.noise %>% 
#   str_subset(pattern = "03|06|12|18|21")

(dom.plot.ME.noise <- P.long.noise.df %>% 
    magrittr::extract(dense.names.noise) %>% 
    map_dfr(~dplyr::select(., cell, cell.kind, tile.id.num, s.mismatch.mag, noise.level)) %>% 
    filter(cell == ("ME.100.C.2")) %>% 
    mutate(noise.level = factor(noise.level, levels = order.levels.noise)) %>% 
    # filter(noise.level %in% paper.filter.noise) %>% 
    left_join(area$area.sf, by = "tile.id.num") %>% 
    st_as_sf() %>% 
    ggplot() +
    geom_sf(aes(fill = s.mismatch.mag), color = "transparent") +
    scale_fill_viridis(na.value = "transparent") +
    facet_wrap(vars(noise.level), ncol = 4) +
    coord_sf() +
    labs(
      title = "Meso noise on the dominance scale",
      color = "Signal\ndominance", fill = "Signal\ndominance",
      x = "", y = "") +
    theme_bw() +
    theme(axis.text.x = element_blank(),
          axis.text.y = element_blank(),
          axis.ticks.x = element_blank(),
          axis.ticks.y = element_blank(),
          legend.position = "bottom",
          rect = element_rect(fill = "transparent")))
ggsave("workflow_objects/noise.dom.plot.ME.w.th.png", dom.plot.ME.noise, device = "png", bg = "transparent",
       height = 5)
# ggsave("Final Plots/noise.dom.plot.ME.eps", dom.plot.ME, device = "eps")
```

### Quantization

```{r}
# define the number of bits --> final number of levels are 2^n + 1

quant.levels <- c(
  quant.00 = 0,
  quant.01 = 1,
  quant.02 = 2,
  quant.03 = 3,
  quant.04 = 4,
  quant.05 = 5,
  quant.10 = 10,
  quant.true = 0) # last version is just a placeholder, the "0" is meaningless

quant.levels.wo.true <- names(quant.levels)[!names(quant.levels) == ("quant.true")]
quant.levels.only.true <- names(quant.levels)[names(quant.levels) == ("quant.true")]



quant_mutate <- function(df, n) {
  mutate(df, "s.quant.{n}" := quantize_mag(x = s, n = n))
}


P.long.quant.df.helper <- cellplans.list %>% 
  map(~mutate(.x$signal.strength.comb.dt, cell.kind = substr(cell, 1, 2))) %>% 
  map2(., cellplans.list, ~left_join(.x, .y$param.df, by = "cell.kind")) %>% 
  map_dfr(~dplyr::rename(., sig_d_th = dominance.th), .id = "network.kind") %>% 
  group_by(network.kind, cell.kind) 

P.long.quant.complete.df <- quant.levels %>% 
  map(~quant_mutate(df = P.long.quant.df.helper, n = .x)) %>% 
  map(ungroup) %>% 
  map_at(quant.levels.only.true, ~mutate(., s.quant.true = s)) %>% 
  map_at(quant.levels.only.true, ~dplyr::select(., -s.quant.0)) %>% 
  map2(., names(quant.levels), ~rename(.x, !!.y := starts_with("s.quant."))) %>% 
  map_at(quant.levels.wo.true, ~dplyr::select(., starts_with("quant."))) %>%
  bind_cols() %>% 
  pivot_longer(cols = starts_with("quant."), names_to = "quant.level", values_to = "quant.mag") %>% 
  split(list(.$network.kind, .$quant.level))
  
names.network.quant <- names(P.long.quant.complete.df)

P.long.quant.df <- P.long.quant.complete.df %>% 
    map(~dplyr::select(., network.kind, matches("tile.id|cell"), quant.level, quant.mag, sig_d_th)) %>% 
  map(~left_join(., area$area.prior, by = "tile.id.chr")) %>% 
  map(~filter(., !quant.mag < 0.05)) %>% # define minimum threshold 
  map(~mutate(., pij = con_llh_sens_custom(., "quant.mag", digits = 3))) 

# saveRDS(P.long.quant.df, "workflow_objects/P.long.quant.df.rds")  
P.long.quant.df <- readRDS("results/P.long.quant.df.rds")

```

```{r}
# quantization on the dominance scale without threshold
order.levels.quant <- c(quant.levels.only.true, rev(quant.levels.wo.true))
dense.names.quant <- names(P.long.quant.df) %>% 
  str_subset("cellplan.3") 

(dom.plot.ME <- P.long.quant.complete.df %>% 
    magrittr::extract(dense.names.quant) %>% 
    map_dfr(~dplyr::select(., cell, cell.kind, tile.id.num, quant.mag, quant.level)) %>% 
    filter(cell == ("ME.100.C.2")) %>% 
    mutate(quant.level = factor(quant.level, levels = order.levels.quant)) %>% 
    left_join(area$area.sf, by = "tile.id.num") %>% 
    st_as_sf() %>% 
    ggplot() +
    geom_sf(aes(fill = quant.mag), color = "transparent") +
    scale_fill_viridis(na.value = "transparent") +
    facet_wrap(vars(quant.level), ncol = 3) +
    coord_sf() +
    labs(title = "Meso quantization on the dominance scale without minimum threshold", 
         color = "Signal\ndominance", fill = "Signal\ndominance", x = "", y = "") +
    theme(axis.text.x = element_blank(),
          axis.text.y = element_blank(),
          axis.ticks.x = element_blank(),
          axis.ticks.y = element_blank(),
          legend.position = "bottom"))
# ggsave("workflow_objects/quant.dom.plot.ME.wo.th.png", dom.plot.ME, device = "png")
```

```{r}
# quantization on the dominance scale with threshold

# paper.filter.quant <- order.levels.quant %>% 
#   str_subset(pattern = "10|04|03|01|00")

(dom.plot.ME.th.quant <- P.long.quant.df %>% 
    magrittr::extract(dense.names.quant) %>% 
    map_dfr(~dplyr::select(., cell, cell.kind, tile.id.num, quant.mag, quant.level)) %>% 
    filter(cell == ("ME.100.C.2")) %>%
    mutate(quant.level = factor(quant.level, levels = order.levels.quant)) %>% 
    # filter(quant.level %in% paper.filter.quant) %>% 
    left_join(area$area.sf, by = "tile.id.num") %>% 
    st_as_sf() %>% 
    ggplot() +
    geom_sf(aes(fill = quant.mag), color = "transparent") +
    scale_fill_viridis(na.value = "transparent") +
    facet_wrap(vars(quant.level), ncol = 4) +
    coord_sf() +
    labs(
      title = "Meso quantization on the dominance scale",
      color = "Signal\ndominance", fill = "Signal\ndominance",
         x = "", y = "") +
    theme_bw() +
    theme(axis.text.x = element_blank(),
          axis.text.y = element_blank(),
          axis.ticks.x = element_blank(),
          axis.ticks.y = element_blank(),
          legend.position = "bottom",
          rect = element_rect(fill = "transparent")))
ggsave("workflow_objects/quant.dom.plot.ME.w.th.png", dom.plot.ME.th.quant, device = "png", bg = "transparent",
       height = 5)
# ggsave("Final Plots/quant.dom.plot.ME.w.th.eps", dom.plot.ME.th, device = "eps")
```

```{r}
cell.kind.unique <- n_distinct(cellplans.list$cellplan.1$param.df$cell.kind)
range.total <- rep(seq(10, 20000, by = 10), cell.kind.unique)
length.range.total <- rep(length(range.total) / cell.kind.unique, cell.kind.unique)

# use the helpers to construct plot dataframe and join with the input params
df <- tibble(cell.kind = factor(rep(cellplans.list$cellplan.1$param.df$cell.kind, length.range.total)),
             distance = range.total) %>%
  left_join(cellplans.list$cellplan.1$param.df, by = "cell.kind") %>%
  mutate(dBm = W2dBm(W)) %>%
  mutate(distance.log10 = log10(distance)) %>%
  mutate(dBm = distance2dB(distance, ple, W)) %>%
  mutate(s = db2s(dBm,
                  midpoint = midpoint,
                  steepness = steepness))

### theoretical quantization vs true
new.df.theo.vars <- quant.levels %>%
  map(~quant_mutate(df = df, n = .x)) %>%
  map_at(quant.levels.only.true, ~mutate(., s.quant.true = s)) %>%
  map_at(quant.levels.only.true, ~dplyr::select(., -s.quant.0)) %>%
  map2(., names(quant.levels), ~rename(.x, !!.y := starts_with("s.quant."))) %>%
  map_at(quant.levels.wo.true, ~dplyr::select(., starts_with("quant."))) %>%
  bind_cols() %>%
  pivot_longer(cols = starts_with("quant."), names_to = "quant.level", values_to = "quant.mag") %>%
  mutate(below.dominance.th = case_when(quant.mag >= 0.05 ~ "Above",
                                        quant.mag < 0.05 ~ "Below"))


(theo.quant.plot.new <- new.df.theo.vars %>%
    mutate(quant.level = factor(quant.level, levels = order.levels.quant)) %>% 
  # filter(!quant.level %in% c("quant.04", "quant.05")) %>%
  ggplot() +
  # geom_point(aes(x = s, y = quant.mag, color = quant.level, group = quant.level))
  geom_line(aes(x = s, y = quant.mag,
                color = quant.level),
            size = 1) +
  facet_wrap(vars(quant.level), ncol = 4) +
  labs(x = "True signal dominance",
       y = "Quantized signal dominance",
       color = "Quant. Version") +
  theme_bw() +
  theme(legend.position = "bottom"))
ggsave("workflow_objects/theo.quant.plot.new.png", theo.quant.plot.new, device = "png")
```

## Estimation 

### Preparing the parameter lists

#### Probabilistic Estimators

```{r estimation-input-parameters}
P.long.mismatch.complete <- c(P.long.noise.df, P.long.quant.df) %>%
  discard(str_detect(names(.), "quant.true")) 
mismatch.versions <- names(P.long.mismatch.complete)

P.long.names <- c(names(noise.expression), names(quant.levels)) %>% 
  str_subset("quant.true", negate = T)


# Each toyworld produces a separate c.vector, however it stays consistent within each toyworld.


### C vector, adding antennas that have 0 phones to complete the vector, arranging it according to the antenna.ID and saving as vector
c.vec <- gen.model.objects$P.long.complete.df %>%
  map(~distinct(., cell.num, phones.sum)) %>%
  map2(., cellplans.list, ~right_join(.x, .y$cellplan.combined.df, by = "cell.num")) %>%
  map(~mutate(., phones.sum = case_when(is.na(phones.sum) ~ 0,
                                TRUE ~ phones.sum))) %>%
  map(~arrange(., cell.num))


# Number of priors
n.prior <- names(area$area.df) %>% 
  str_count("prior") %>% 
  sum()
  

### final list objects

# P.star
P.long.mismatch.parameter.list <- P.long.mismatch.complete %>% 
  rep(n.prior) %>% 
  .[order(names(.))] # order alphabetically because it also has toyworld name in

# c.vec
c.vec.parameter.list <- c.vec %>%
  map(~data.table(i = .x$cell.num, c = .x$phones.sum)) %>% 
  rep(length(P.long.names) * n.prior) %>% 
  .[order(names(.))] # order alphabetically as above

# prior name
prior.parameter.list <- names(area$area.df) %>% 
  str_subset("prior") %>% 
  rep(length(mismatch.versions)) %>% 
  as.list() #dont order alphabetically as it should alternate


prob.est.input.names <- cross(list(unique(names(c.vec.parameter.list)),
                                   P.long.names,
                                   unique(prior.parameter.list))) %>% 
  map(lift(paste, sep = "_")) %>% 
  set_names(unlist(.)) %>% 
  .[order(names(.))] # order alphabetically 

# parameter list for probability based estimators
prob.est.input.list <- list(c = c.vec.parameter.list,
                            P = P.long.mismatch.parameter.list,
                            a = prior.parameter.list) %>% 
  map(~set_names(., prob.est.input.names)) ### is this the correct order (qm))))


# number of iterations
n.iter.MLE = 200
```


#### Determnistic Parameters

```{r estimation-input-parameters}

cellplan.combined.parameter.list <- flatten(cellplans.list) %>% 
  keep(str_detect(names(.), "cellplan.combined.df")) %>%
  set_names(names(cellplans.list)) %>% 
  rep(n.prior) %>% 
  .[order(names(.))] # order alphabetically as above

signal.strength.comb.dt.parameter.list <- flatten(cellplans.list) %>% 
  keep(str_detect(names(.), "signal.strength.comb.dt")) %>% 
  set_names(names(cellplans.list)) %>% 
  rep(n.prior) %>% 
  .[order(names(.))] # order alphabetically as above

C.vec.df.parameter.list <- gen.model.objects$C.vec.df.list %>% 
  rep(n.prior) %>% 
  .[order(names(.))] # order alphabetically as above


prior.var.parameter.list <- names(area$area.df) %>% 
  str_subset("prior") %>% 
  rep(length(names(cellplans.list))) %>% 
  as.list() #dont order alphabetically as it should alternate

det.est.input.names <- cross(list(unique(names(C.vec.df.parameter.list)),
                                   unique(prior.var.parameter.list))) %>% 
  map(lift(paste, sep = "_")) %>% 
  set_names(unlist(.)) %>% 
  .[order(names(.))] # order alphabetically 

# parameter list for deterministic estimators
VOR.input.list <- list(cellplan.combined = cellplan.combined.parameter.list, 
                       signal.strength.comb.dt = signal.strength.comb.dt.parameter.list, 
                       C.vec.df = C.vec.df.parameter.list,
                       prior.var = prior.var.parameter.list) %>% 
  map(~set_names(., det.est.input.names))
```




### Voronoi estimators

```{r VOR.tower, warning=FALSE, echo=FALSE, out.width="50%"}
# Voronoi estimation with tower locations as seeds
VOR.tower.names.est <- paste0("u.VOR.tower_", det.est.input.names)
VOR.tower <- pmap(VOR.input.list, ~VOR_est(area = area, 
                                           cellplan.combined = ..1, 
                                           signal.strength.comb.dt = ..2, 
                                           C.vec.df = ..3, 
                                           prior.var = ..4,
                                           seed = "tower"))

VOR.tower.est <- VOR.tower %>% 
  map(~mutate(.x$seed.voronoi.final, tile.id.chr = as.character(tile.id))) %>% 
  map_at(c(1), ~dplyr::select(., tile.id.chr, u.VOR)) %>% 
  map_at(-c(1), ~dplyr::select(., u.VOR)) %>% 
  map2(., VOR.tower.names.est, ~rename_with(.x, 
                                            stringr::str_replace, 
                                            pattern = "u.VOR", replacement = .y, 
                                            .cols = starts_with("u.VOR"))) %>% 
  bind_cols()
```

```{r VOR.offset, warning=FALSE, echo=FALSE, out.width="50%"}
# Voronoi estimation with cell locations + offset as seeds
VOR.offset.names.est <- paste0("u.VOR.offset_", det.est.input.names)
VOR.offset <- pmap(VOR.input.list, ~VOR_est(area = area, 
                                            cellplan.combined = ..1, 
                                            signal.strength.comb.dt = ..2, 
                                            C.vec.df = ..3, 
                                            prior.var = ..4,
                                            seed = "cell.offset",
                                            offset = 10)) 

VOR.offset.est <- VOR.offset %>% 
  map(~mutate(.x$seed.voronoi.final, tile.id.chr = as.character(tile.id))) %>% 
  map_at(c(1), ~dplyr::select(., tile.id.chr, u.VOR)) %>% 
  map_at(-c(1), ~dplyr::select(., u.VOR)) %>% 
  map2(., VOR.offset.names.est, ~rename_with(.x, 
                                            stringr::str_replace, 
                                            pattern = "u.VOR", replacement = .y, 
                                            .cols = starts_with("u.VOR"))) %>% 
  bind_cols()
```

```{r VOR.barycenter, warning=FALSE, echo=FALSE, out.width="50%"}
# Voronoi estimation with cell barycenter locations as seeds
VOR.barycenter.names.est <- paste0("u.VOR.barycenter_", det.est.input.names)
VOR.barycenter <- pmap(VOR.input.list, ~VOR_est(area = area, 
                                                cellplan.combined = ..1, 
                                                signal.strength.comb.dt = ..2, 
                                                C.vec.df = ..3, 
                                                prior.var = ..4,
                                                seed = "cell.barycenter")) 

VOR.barycenter.est <- VOR.barycenter %>% 
  map(~mutate(.x$seed.voronoi.final, tile.id.chr = as.character(tile.id))) %>% 
  map_at(c(1), ~dplyr::select(., tile.id.chr, u.VOR)) %>% 
  map_at(-c(1), ~dplyr::select(., u.VOR)) %>% 
  map2(., VOR.barycenter.names.est, ~rename_with(.x, 
                                            stringr::str_replace, 
                                            pattern = "u.VOR", replacement = .y, 
                                            .cols = starts_with("u.VOR"))) %>% 
  bind_cols()

```

### MLE/EM estimator

```{r MLE-estimator}

MLE.names.est <- paste0("u.MLE_", prob.est.input.names)

MLE.est <- pmap(prob.est.input.list,
                 ~EM_est_supertiles(area = area,
                                    c.vec.dt = ..1, 
                                    P.dt = ..2, 
                                    prior.var = ..3,
                                    selected.range = c(1, 10, 50, 200),
                                    n.iter = n.iter.MLE,
                                    message = T, 
                                    ldt.dt = 0)) %>% 
  map(~arrange(., tile.id.num)) %>% 
  map_at(c(1), ~dplyr::select(., tile.id.num, starts_with("u_"))) %>% 
  map_at(-c(1), ~dplyr::select(., starts_with("u_"))) %>% 
  map2(., MLE.names.est, ~rename_with(.x, 
                                             stringr::str_replace, 
                                             pattern = "u", replacement = .y, 
                                             .cols = starts_with("u"))) %>% 
  bind_cols()
```

### DF estimator

```{r DF-estimator}
DF.names.est <- paste0("u.DF_", prob.est.input.names)
## Renormalizing with EM and bringing estimate on regular tile.id level
DF.est <- pmap(prob.est.input.list,
                ~DF_est_relaxed_iter_supertiles(area = area,
                                                c.vec.dt = ..1, 
                                                P.dt = ..2, 
                                                prior.var = ..3,
                                                selected.range = c(1, 10, 50, 200),
                                                n.iter = n.iter.MLE,
                                                message = T, 
                                                ldt.dt = 0)) %>% 
  map(~arrange(., tile.id.num)) %>% 
  map_at(c(1), ~dplyr::select(., tile.id.num, starts_with("u_"))) %>% 
  map_at(-c(1), ~dplyr::select(., starts_with("u_"))) %>% 
                   map2(., DF.names.est, ~rename_with(.x, 
                                      stringr::str_replace, 
                                      pattern = "u", replacement = .y, 
                                      .cols = starts_with("u"))) %>% 
  bind_cols()
```

```{r estimation-combine}
### Evaluation
final.estimates.sf <- area$area.sf %>% 
  left_join(VOR.tower.est, by = "tile.id.chr") %>%
  left_join(VOR.offset.est, by = "tile.id.chr") %>%
  left_join(VOR.barycenter.est, by = "tile.id.chr") %>%
  left_join(MLE.est, by = "tile.id.num") %>%
  left_join(DF.est, by = "tile.id.num") %>%
  arrange(tile.id.num)
  # group_by(prior.2) %>%
  # mutate(u.grouped.flat = mean(pop)) %>%
  # ungroup() %>%
  # mutate(u.flat = mean(pop))

saveRDS(final.estimates.sf, "workflow_objects/final.estimates.sf.rds")


# non-sf version
final.estimates.df <- final.estimates.sf %>% 
  st_drop_geometry()

# vector with names of the relevant estimates
names.final.estimates <- final.estimates.sf %>% 
  dplyr::select(pop, starts_with("u.")) %>% # all estimates
  # dplyr::select(pop, matches(as.character(n.iter.MLE)), matches("VOR")) %>% # only "final" ones for mapping
  st_drop_geometry() %>% 
  names()
```

## Evaluation

### Spatial Density

```{r estimation-maps}
# Define break points for discretized spatial density plots
breaks <- c(0, 2, 5, 10, 20, 50, 100, 200, 350, Inf)
maps.input <- final.estimates.sf %>% 
  dplyr::select(tile.id, pop, X.centroid, Y.centroid, all_of(names.final.estimates)) %>% 
  mutate(across(c(pop, starts_with("u.")), ~cut(., breaks = breaks, dig.lab = 7, right = F)))

# Build maps and print
(maps.estimation.density <- names.final.estimates %>%
    map(~map_density(data = maps.input, var = .x, label = .x)) %>%
    set_names(names.final.estimates))

# saving maps
# walk2(maps.estimation.density, names(maps.estimation.density),
#       ~ggsave(filename = paste0(.y, ".png"), plot = .x,
#               path = paste0(getwd(), "/Poster/"), device = "png"))


```

### KWD all estimators

```{r kwd-computation}
# develop dataframe with GTP, all final estimates and tile centroids
kwd.helper.est <- final.estimates.sf %>% 
  dplyr::select(tile.id.num, pop, all_of(names.final.estimates)) %>%
  # dplyr::select(-c(elevation, type), -matches("prior")) %>%
  st_centroid() %>% 
  mutate(lon = unlist(map(.$geometry, 1)),
         lat = unlist(map(.$geometry, 2))) %>% 
  st_drop_geometry()

# Coordinates object
coordinates <- kwd.helper.est %>% 
  dplyr::select(lon, lat) %>% 
  as.matrix()

# Weights object
weights <- kwd.helper.est %>% 
  # dplyr::select(pop, matches("200")) %>%
  dplyr::select(pop, starts_with("u.")) %>%
  as.matrix()


# Approximation parameter (the higher the more accurate)
L = 3

# Run KWD
kwd.final <- compareOneToMany(coordinates, weights, L = L, recode = TRUE)
paste("KWD runtime ( L =", L, "):", round(kwd.final$runtime / 60, 0), "min for", 
      ncol(weights) - 1, "estimates and", 
      length(final.estimates.df$tile.id), "tiles")

# Define names for estimators
names.weights <- colnames(weights)[-1]

order.mismatch <- str_replace(mismatch.versions, paste0(names(gen.model.objects$C.vec.df.list), "."), "")


# Develop data frame on the estimtor level with respective KWD values
kwd.eval.caro <- tibble(estimator = names.weights,
                   kwd = kwd.final$distance * 1) %>%  # rescaling according to scale
  mutate(kwd.lower.bound = kwd - ((kwd / 100) * 1.29)) %>%  # for L=3 within 1 percent 
  mutate(network.kind = str_extract(estimator, pattern = paste0(names(gen.model.objects$C.vec.df.list), collapse = "|"))) %>% 
  mutate(iteration = str_extract(estimator, "[[:digit:]]+$")) %>% 
  mutate(version = case_when(str_detect(estimator, "true") ~ "true",
                             TRUE ~ str_extract(estimator, 
                                                pattern = paste(c(order.mismatch, "flat"), collapse = "|")))) %>% 
  # kind = substr(estimator, 1, 11)) %>%
  mutate(iteration = case_when(is.na(iteration) ~ 0,
                               TRUE ~ as.numeric(iteration))) %>% 
  mutate(estimator.kind = str_extract(estimator, pattern = "MLE|DF|flat")) %>% 
  group_by(version, estimator.kind) %>% 
  mutate(final.iteration = case_when(iteration == max(iteration) ~ T,
                                     T ~ F)) %>% 
  mutate(min.KWD = case_when(kwd == min(kwd) ~ T,
                             T ~ F)) %>% 
  ungroup()  %>% 
  mutate(version.kind = case_when(str_detect(version, "no") ~ "Noise",
                                  str_detect(version, "quant") ~ "Quantization",
                                  str_detect(version, "true") ~ "True",
                                  is.na(version) ~ "True")) %>% 
    mutate(prior.version = case_when(str_detect(estimator, "prior.1") ~ "Uninformative",
                                   str_detect(estimator, "prior.2") ~ "Mildly informative")) %>% 
  mutate(estimator.name = case_when_estimator(estimator, iteration)) %>% 
  group_by(version, estimator.name) %>% 
  mutate(final.iteration = case_when(iteration == max(iteration) ~ T,
                                     T ~ F)) %>% 
  mutate(min.KWD = case_when(kwd == min(kwd) ~ T,
                             T ~ F)) %>% 
  ungroup()

saveRDS(kwd.eval, "workflow_objects/kwd.eval.rds")
kwd.eval.caro <- readRDS("results/kwd.eval.rds")
order.mismatch <- readRDS("results/results/order.mismatch.caro.rds")

case_when_estimator <- function(estimator, iteration) {
  case_when(
    str_detect(estimator, "u.VOR.tower") ~ "VOR.t",
    str_detect(estimator, "u.VOR.offset") ~ "VOR.o",
    str_detect(estimator, "u.VOR.barycenter") ~ "VOR.b",
    str_detect(estimator, "u.MLE") & iteration == 1 ~ "SB",
    str_detect(estimator, "u.MLE") & iteration == 10 ~ "MLE10",
    str_detect(estimator, "u.MLE") ~ "MLE200",
    str_detect(estimator, "u.DF") & iteration == 1 ~ "DF1",
    str_detect(estimator, "u.DF") & iteration == 10 ~ "DF10",
    str_detect(estimator, "u.DF") ~ "DF200",
    str_detect(estimator, "u.grouped.flat") ~ "Grouped.flat",
    str_detect(estimator, "u.flat") ~ "flat",
    
  )
}


kwd.eval <- kwd.eval.caro %>% 
  mutate(version.kind = case_when(str_detect(version, "no") ~ "Noise",
                                  str_detect(version, "quant") ~ "Quantization",
                                  str_detect(version, "true") ~ "True",
                                  is.na(version) ~ "True")) %>% 
  mutate(prior.version = case_when(str_detect(estimator, "prior.1") ~ "Uninformative",
                                   str_detect(estimator, "prior.2") ~ "Mildly informative")) %>% 
  mutate(estimator.name = case_when_estimator(estimator, iteration)) %>% 
  group_by(version, estimator.name) %>% 
  mutate(final.iteration = case_when(iteration == max(iteration) ~ T,
                                     T ~ F)) %>% 
  mutate(min.KWD = case_when(kwd == min(kwd) ~ T,
                             T ~ F)) %>% 
  ungroup()
```

### KWD for estimates with no model mismatch

```{r kwd_1}

difference <- kwd.eval %>% 
  filter(str_detect(estimator, "VOR|flat") | final.iteration == T) %>%
  filter(version.kind == "True") %>%
  mutate(estimator.name.ordered = reorder(estimator.name, -kwd)) %>%
  group_by(network.kind, estimator.name.ordered) %>% 
  summarise(diff.per = ((max(kwd) - min(kwd)) / max(kwd)) * -100,
            y.axis.brace = max(kwd) + 1.4,
            y.axis.dif = max(kwd) + 1.7)
  
  

(kwd.final.estimates.plot <- kwd.eval %>% 
    filter(str_detect(estimator, "VOR|flat") | final.iteration == T) %>%
    filter(version.kind == "True") %>% 
    mutate(estimator.name.ordered = reorder(estimator.name, -kwd)) %>%
    ggplot(aes(x = estimator.name.ordered, y = kwd)) + 
    geom_bar(aes(fill = prior.version), 
             stat = "identity", position = position_dodge(width = 0.9)) + 
    geom_errorbar(aes(ymin = kwd.lower.bound, ymax = kwd, fill = prior.version), 
                  position = position_dodge(width = 0.9), width = 0.25) +
    geom_text(aes(label = round(kwd, 2), fill = prior.version), 
              position = position_dodge(1), hjust = -0.1, color = "Black", size = 3) +
    geom_text(data = difference, aes(x = estimator.name.ordered, y = y.axis.brace, 
                                     label = paste("}")),
              position = position_dodge(1), hjust = 0, color = "Blue", size = 4) +
    geom_text(data = difference, aes(x = estimator.name.ordered, y = y.axis.dif, 
                                     label = paste(round(diff.per, 2), "%")),
              position = position_dodge(0.5), hjust = -0.1, color = "Blue", size = 4) +
    scale_fill_ptol() +
    ylim(0, 15) +
    facet_wrap(vars(network.kind)) +
    theme_bw() +
    coord_flip() +
    labs(title = "KWD for estimates with no model mismatch",
         x = "", y = "KWD", 
         alpha = "Iteration", fill = "Prior") + 
    theme(legend.position = "bottom"))

ggsave("kwd.evaluation.with.informative.prior.png", kwd.final.estimates.plot, dev = "png")
```

### KWD after `r n.iter.MLE` iterations

```{r kwd-noise}
order.noise.levels <- c(str_subset(order.mismatch, "true"), 
                        str_subset(order.mismatch, "true|quant", negate = T)) %>%
  str_replace(., "no.true", "true")  %>% 
  unique(.)

order.quant.levels <- c(str_subset(order.mismatch, "true"),
                        rev(str_subset(order.mismatch, "no.|true", negate = T))) %>% 
  str_replace(., "no.true", "true")  %>% 
  unique(.)

mismatch.helper <- c(order.noise.levels, order.quant.levels[-1])

kwd.vor.helper <- kwd.eval %>% 
  mutate(estimator.name = case_when_estimator(estimator, iteration)) %>% 
  filter(str_detect(estimator.name, "VOR")) %>% 
  mutate(time = case_when(str_detect(estimator.name, "VOR") ~ 15,
                          TRUE ~ 1)) %>% 
  uncount(time) %>% 
  mutate(version = rep(mismatch.helper, 24))
  
  
(kwd.noise.plot <- kwd.eval %>% 
    filter(estimator.name == "SB" | iteration == 10) %>% 
    filter(version.kind %in% c("Noise", "True")) %>% 
    bind_rows(filter(kwd.vor.helper, str_detect(version, "no.|true"))) %>% 
    mutate(estimator.name.ordered = reorder(estimator.name, kwd)) %>%
    mutate(version.ordered = factor(version, levels = order.noise.levels)) %>%
    ggplot(aes(x = version.ordered, y = kwd, color = estimator.name.ordered, group = estimator.name.ordered)) +
    geom_line(size = 1) +
    geom_point() +
    geom_ribbon(aes(ymin = kwd.lower.bound, ymax = kwd, group = version), 
                alpha = 0.2, color = "transparent") +
    # geom_text_repel(aes(label = round(kwd, 2)), size = 7, show.legend  = F) +
    colorspace::scale_fill_discrete_qualitative(palette = "Harmonic") +
    facet_wrap(vars(network.kind, prior.version), scales = "free_x", ncol = 4) +
    theme_bw() +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1),
          legend.position = "bottom") +
    labs(x = "Version",
         y = "KWD", color = "Estimator",
         title = "KWD estimates random noise model mismatch (after 10 iterations)"))

ggsave("kwd.noise.plot.10.png", kwd.noise.plot, dev = "png")
  


(kwd.quant.plot <- kwd.eval %>% 
    filter(estimator.name == "SB" | iteration == 1) %>% 
    filter(version.kind %in% c("Quantization", "True")) %>% 
    bind_rows(filter(kwd.vor.helper, str_detect(version, "quant.|true"))) %>% 
    mutate(estimator.name.ordered = reorder(estimator.name, kwd)) %>%
    mutate(version.ordered = factor(version, levels = order.quant.levels)) %>%
    ggplot(aes(x = version.ordered, y = kwd, color = estimator.name.ordered, group = estimator.name.ordered)) +
    geom_line(size = 1) +
    geom_point() +
    geom_ribbon(aes(ymin = kwd.lower.bound, ymax = kwd, group = version), 
                alpha = 0.2, color = "transparent") +
    # geom_text_repel(aes(label = round(kwd, 2)), size = 7, show.legend  = F) +
    colorspace::scale_fill_discrete_qualitative(palette = "Harmonic") +
    facet_wrap(vars(network.kind, prior.version), scales = "free_x", ncol = 4) +
    theme_bw() +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1),
          legend.position = "bottom") +
    labs(x = "Version",
         y = "KWD", color = "Estimator",
         title = "KWD estimates quantization model mismatch (after 10 iterations)"))

ggsave("kwd.quant.plot.10.png", kwd.quant.plot, dev = "png")
```

### KWD convergence for numerical estimators

```{r kwd-convergence, warning=FALSE}
# log scale breaks for background grid
minor.breaks <- rep(1:9, 21) * (10^rep(-10:10, each = 9))

# develop line plot
(kwd.convergence.noise.plot <- kwd.eval %>% 
    # filter(str_detect(estimator, "prior.1")) %>%
    filter(!estimator.kind == "flat") %>% 
    filter(version.kind %in% c("Noise", "True")) %>% 
    mutate(version.ordered = factor(version, levels = order.noise.levels)) %>%
    ggplot(aes(x = iteration, y = kwd, color = version.ordered)) +
    geom_line() +
    geom_point(show.legend = F) +
        # geom_text_repel(aes(label = round(kwd, 2)), size = 3, show.legend  = F) +
    facet_wrap(vars(network.kind, estimator.kind, prior.version), nrow = 2) +
    geom_ribbon(aes(ymin = kwd.lower.bound, ymax = kwd, group = version.ordered), 
                alpha = 0.2, color = "transparent") +
    scale_color_ptol() +
    scale_x_log10(labels = scales::trans_format("log10", scales::math_format(10^.x)),
                  minor_breaks = minor.breaks) +
    annotation_logticks(sides = "b") +
    labs(color = "Random Noise",
         x = "Iteration",
         y = "KWD",
         title = "Random Noise: KWD convergence behavior per estimator"))

ggsave("kwd.convergence.noise.plot.png", kwd.convergence.noise.plot, dev = "png",
       width = 10)

(kwd.convergence.quant.plot <- kwd.eval %>% 
    # filter(str_detect(estimator, "prior.1")) %>%
    filter(!estimator.kind == "flat") %>% 
    filter(version.kind %in% c("Quantization", "True")) %>% 
    mutate(version.ordered = factor(version, levels = order.quant.levels)) %>%
    ggplot(aes(x = iteration, y = kwd, color = version.ordered)) +
    geom_line() +
    geom_point(show.legend = F) +
        # geom_text_repel(aes(label = round(kwd, 2)), size = 3, show.legend  = F) +
    facet_wrap(vars(network.kind, estimator.kind, prior.version), nrow = 2) +
    geom_ribbon(aes(ymin = kwd.lower.bound, ymax = kwd, group = version.ordered), 
                alpha = 0.2, color = "transparent") +
    scale_color_ptol() +
    scale_x_log10(labels = scales::trans_format("log10", scales::math_format(10^.x)),
                  minor_breaks = minor.breaks) +
    annotation_logticks(sides = "b") +
    labs(color = "Quantization",
         x = "Iteration",
         y = "KWD",
         title = "Quantization: KWD convergence behavior per estimator"))

ggsave("kwd.convergence.quant.plot.png", kwd.convergence.quant.plot, dev = "png",
       width = 10)

```




```{r}
sessionInfo()
```


